EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-05-12 04_cbinding.tex
First complete draft of the edipack2triqs implementation subsection

\subsection{TRIQS interface}\label{sSecInteropTRIQS}
{\color{red} Igor }
A purely Python \NAME to Triqs interface is available, leveraging on
the C-bindings and Python API. The corresponding module depends on
\NAME (which ultimately depends on SciFortran) and Triqs.
Assuming the two software are correctly installed in the OS, the
EDIpack2Triqs interface is installed as follows:

@@

@@

The c-bindings approach is extremely handy, in that is opens the way for interoperability 
with a large number of languages and frameworks. As an example, an experimental Julia API 
is provided in the EDIpack github repository. Rather than a fully functioning Julia package, 
this is at present a proof of concept, capable of replicating the results of Fortran, C++ and
python implementations for the Bethe lattice example driver.
Its structure and operation mimicks that of the EDIpy layer, with minimal language-specific differences.

The EDIpack Julia api consists of a module called {\tt EDIjl2}, which provides access to the global variables and functions contained in {\tt  libedipack2\_cbinding.so/.dylib}. This is searched upon loading the module, with the following loading hierarchy:

\begin{itemize}
\item {\tt EDIPACK\_PATH}: if this environment variable is set, {\tt EDIjl} will look there for the library first
\item {\tt LD\_LIBRARY\_PATH}: for Linux systems
\item {\tt DYLD\_LIBRARY\_PATH}: for Mac systems
\end{itemize}

in complete analogy to the Python API, the global variables and functions are contained in a  {\tt struct} called {\tt global\_env}, and can be called and assigned in the following way:

\begin{itemize}
\item A variable symbol is retrieved from the library through {\tt Libdl.dlsym} and then read via {\tt unsafe\_load}
\item Similarly, variable assignment is achieved through {\tt Libdl.dlsym} and  {\tt unsafe\_store}
\item Functions are called via {\tt Libdl.dlsym} and  {\tt ccall}
\end{itemize}

\end{document}

>>>>

\subsection{TRIQS interface}\label{sSecInteropTRIQS}
A thin compatibility layer between \NAME and the Toolbox for Research on
Interacting Quantum Systems (TRIQS)\cite{Parcollet2015CPC} called edipack2triqs
is available as a stand-alone project. It is a pure Python package
built upon \NAME's Python API  (Sec.~\ref{sSecInteropEDIpy}) that provides a
limited object-oriented interface to the most important features of \NAME.
edipack2triqs strives to offer seamless interoperability with program tools
based on TRIQS by adopting data types, conventions and usage patterns
common to other TRIQS-based impurity solvers, such as
TRIQS/CTHYB\cite{Seth2016CPC}. It also enables execution of \NAME calculations
in the MPI parallel mode with no extra effort from the user.
Detailed documentation of the package can be 
found online at \href{https://krivenko.github.io/edipack2triqs/}
{https://krivenko.github.io/edipack2triqs/}.

\subsubsection{Installation}\label{sSecInteropTRIQSInstallation}
The package depends on \NAME (which ultimately depends on SciFortran), EDIpy2
and TRIQS version 3.1 or newer. Assuming the three prerequisites are correctly 
installed in the OS, the current development version of edipack2triqs can be 
installed with {\tt pip} from its 
\href{https://github.com/krivenko/edipack2triqs}{GitHub repository} as follows:

@@

\paragraph{Anaconda}
Another option for installing the package is by using the Anaconda package 
manager on Unix/Linux systems.
The following commands will create a new {\tt conda} environment named 
`edipack' and install the most recently released version of edipack2triqs along 
with its dependencies (\NAME, EDIpy2 and the TRIQS libraries).

\begin{lstlisting}[style=mybash]
conda create -n edipack
conda activate edipack
conda install -c conda-forge -c edipack edipack2triqs
\end{lstlisting}

\subsubsection{Implementation}\label{sSecInteropTRIQSImplementation}
The programming interface of edipack2triqs is built around the singleton 
Python class {\tt EDIpackSolver} (defined in the module
{\tt edipack2triqs.solver}), whose instance represents the internal state of 
the \NAME library and exposes its functionality through a number of 
attributes and methods.

\paragraph{{\bf Constructor}}
Constructor of {\tt EDIpackSolver} accepts the Hamiltonian to be diagonalized
in the form of a second-quantized fermionic operator, the TRIQS {\tt Operator}
object described in Sec.~8.7 of Ref.~\onlinecite{Parcollet2015CPC} (impurity 
problems involving bosons are not supported in the current version). In 
addition to the Hamiltonian, four {\it fundamental operator sets}
{\tt fops\_imp\_up}, {\tt fops\_imp\_dn}, {\tt fops\_bath\_up} and
{\tt fops\_bath\_dn} must be provided.
Each of the sets contains pairs of labels {\tt (b,j)} carried
by the operators $c^\dagger_{b,i} / c_{b,i}$ corresponding to either impurity
or bath electronic degrees of freedom with a certain spin projection.
Having these two crucial pieces of information, the constructor initializes
the underlying Fortran library and automatically selects the appropriate
{\tt ed\_mode=normal,superc,nonsu2} (Sec.~\ref{sSecQNs}) and
{\tt bath\_type=normal,hybrid,general} (Sec.~\ref{sSecBath}). Besides, the
constructor accepts a vast array of keyword arguments that allow for fine-tuning
the diagonalization process. Among others, these include the ED algorithm
selection, the quantum numbers to be used, parameters of the Krylov space,
the spectrum cutoff and various tolerance levels.

\paragraph{{\bf Method {\tt solve}}}
The method {\tt solve()} of {\tt EDIpackSolver} calls
{\tt edipy2.global\_env.solve()} and therefore performs the bulk of
calculations. It accepts the inverse temperature $\beta$ required to calculate
the expectation value of physical observables, and parameters of energy grids
for Green's function calculations.

\paragraph{{\bf Input parameters}}
It is possible to read off and change parameters of the Hamiltonian between
successive calls to {\tt solve()} via respective attributes of {\tt 
EDIpackSolver}. Note, however, that the changes that would necessitate a change of {\tt ed\_mode} or {\tt bath\_type} are disallowed. The relevant attributes
are the following.
\begin{itemize}
    \item {\tt nspin} --- the number of non-equivalent spin projections 
          (read-only).
    \item {\tt norb} -- the number of impurity orbitals (read-only).
    \item {\tt hloc} --- matrix $h^0_{\alpha\beta\s\s'}$ of
          the quadratic impurity Hamiltonian (\ref{Himp}).
    \item {\tt Uloc}, {\tt Ust}, {\tt Jh}, {\tt Jx}, {\tt Jp} ---
          parameters $U, U', J, J_X \text{ and } J_P$ of the impurity
          interaction term $\hat{H}^\mathrm{int}_\mathrm{imp}$ (\ref{Hint}).
    \item {\tt bath} --- an object representing the bath. Depending on the
          bath type selected upon solver object construction, this object is an
          instance of {\tt BathNormal}, {\tt BathHybrid} or {\tt BathGeneral}
          (all three classes are members of the module
          {\tt EDIpackSolver.bath}). The bath objects support basic arithmetic
          operations so that a mixing scheme within a DMFT calculation can be
          easily implemented. The way to access individual bath parameters is
          specific to each of the three classes, and is described in detail in the       \href{https://krivenko.github.io/edipack2triqs/documentation.html\#module-edipack2triqs.bath}{online API reference}.
\end{itemize}

\paragraph{{\bf Calculation results}}
After a successful invocation of {\tt solve()}, one can extract results of the 
calculation from the following attributes.
\begin{itemize}
    \item {\tt e\_pot}, {\tt e\_kin} --- thermal average of the potential 
          (interaction) and kinetic energy respectively.
    \item {\tt densities}, {\tt double\_occ} --- lists of average densities and
          double occupancies of impurity orbitals.
    \item {\tt magnetization} --- Cartesian components of the average impurity  
          magnetization vectors, one row per orbital.
    \item {\tt superconductive\_phi} --- Matrix of impurity superconductive
          order parameters
          $\phi_{\alpha\beta} = \langle c_{\alpha\up} c_{\beta\down}\rangle$
          in orbital space. {\color{red} Is it defined elsewhere in the text?}
    \item {\tt g\_iw}, {\tt g\_an\_iw} --- Normal and anomalous components of
          the Matsubara impurity Green's function.
    \item {\tt Sigma\_iw}, {\tt Sigma\_an\_iw} --- Normal and anomalous
          components of the Matsubara impurity self-energy.
    \item {\tt g\_w}, {\tt g\_an\_w} --- Normal and anomalous components of
          the real-frequency impurity Green's function.
    \item {\tt Sigma\_w}, {\tt Sigma\_an\_w} --- Normal and anomalous
          components of the real-frequency impurity self-energy.
\end{itemize}

The Green's functions are returned as TRIQS {\tt BlockGf} containers
with names of the individual blocks determined from the block labels {\tt 'b'}
found in {\tt fops\_imp\_up}, {\tt fops\_imp\_dn} and from solver's
{\tt ed\_mode}. The Matsubara and real-frequency meshes of the TRIQS GF 
containers (see Sec.~8.2 of Ref.~\onlinecite{Parcollet2015CPC}) are constructed
according to the parameters passed to {\tt solve()}.
The anomalous components of the Green's functions and self-energies are only 
available when anomalous bath terms are present in the Hamiltonian.
If this is not the case, an attempt to access these attributes results in a
{\tt RuntimeError}.

\paragraph{{\bf Bath parameter fitting}}
The method {\tt EDIpackSolver.chi2\_fit\_bath()} is essentially a wrapper around
{\tt edipy2.global\_env.chi2\_fitgf()}. It accepts the function to fit (either
the hybridization function or the Weiss field) in the {\tt BlockGf} format,
and returns the parameter fit result as a {\tt Bath*} object along with a
{\tt BlockGf} representation of the fitted function. In the superconducting
case ({\tt ed\_mode=superc}), this method accepts and returns pairs of
the {\tt BlockGf} containers corresponding to the normal and anomalous 
components of the quantities in question.

@@

The c-bindings approach is extremely handy, in that is opens the way for interoperability 
with a large number of languages and frameworks. As an example, an experimental Julia API 
is provided in the EDIpack github repository. Rather than a fully functioning Julia package, 
this is at present a proof of concept, capable of replicating the results of Fortran, C++ and
python implementations for the Bethe lattice example driver.
Its structure and operation mimicks that of the EDIpy2 layer, with minimal language-specific differences.

The \NAME{} Julia api consists of a module called {\tt EDIjl2}, which provides access to the global variables and functions contained in {\tt  libedipack2\_cbinding.so/.dylib}.

The library is searched upon loading the module, with the following order of priority:

\begin{itemize}
\item {\tt EDIPACK\_PATH}: if this environment variable is set, {\tt EDIjl2} will look there for the library first
\item {\tt LD\_LIBRARY\_PATH}: for Linux systems
\item {\tt DYLD\_LIBRARY\_PATH}: for Mac systems
\end{itemize}

\subsubsection{Structure}

In partial analogy to the Python API, the global variables, but not the functions, are contained in a  {\tt struct} called {\tt global\_env}.

Similarly to the C++ case, global variables and functions are loaded as raw pointers. As a consequence, it is important that the dimensions of the array-like variables (such as {\tt ULOC}) and, in general, the amount of memory occupied by each variable are correcty accounted for. This is achieved within the Julia module by appropriately casting the variables to compatible types, such as {\tt  Cint, Cbool, Cdouble}.
Functions are called from the dynamic library, making use of the {\tt  ccall} procedure. As previously stated, as a consequence of the C linking conventions multiple alternative version of the interfaced Fortran procedures are present, to account for the different input variable combinations; the EDIjl2 wrapper functions take care of selecting the appropriate Fortran procedure depending on the set of input parameters provided by the user.

\subsubsection{Installation and usage}

At present, the EDIpack Julia interface is not offered as a package. The git repository has to be cloned via

\begin{lstlisting}[style=mybash]
git clone https://github.com/EDIpack/EDIjl2.0.git
\end{lstlisting}

and the location of the source files needs to be included in the user program via

\begin{lstlisting}[style=mybash]
push!(LOAD_PATH, joinpath(@__DIR__, "PATH/TO/REPO/src"))
\end{lstlisting}

The EDIjl2 module can then be loaded via

\begin{lstlisting}[style=mybash]
using EDIjl2
\end{lstlisting}

The correct way to access global variables and functions is as in the following example:

\begin{lstlisting}[style=mybash]
EDIjl2.read_input("inputED.conf")
ed = EDIjl2.global_env
println("Nspin = ", ed.Nspin)
ed.Nspin = 2
\end{lstlisting}

The names and inputs of the Julia-wrapped functions are entirely analogous to those of the Python API.

Particular care has to be taken with respect to parallelism: Julia offers an in-house platform for parallel computing with memory sharing through multi-threading. At present, there is no integration between this architecture and the MPI-based parallelism of \NAME{}, and their concurrent use in a single program is entirely untested and not encouraged. If the \NAME{} library is compiled with enabled parallelism, MPI support will need to be enabled in the user-created Julia program, by making use of modules such as {\tt  MPI.jl}. The {\tt MPI\_COMM\_WORLD} communicator will need to be correctly initialized by the user as well, and the program will need to be run with {\tt  mpiexec/mpirun}.

We provide an example script for the simple case of the Bethe lattice in the normal phase in the {\tt  examples} folder of the EDIjl2.0 repository. This script is intended to be run serially. Documentation of the API and further examples are being developed.
\end{document}
